---
title: "JRF nat rep demographics"
date: 17/02/2024
format: html
editor: visual
execute: 
  warning: false
  echo: false
---

# Packages

```{r packages}
library(haven)
library(poLCA)
library(dplyr)
library(ggplot2)
library(tidyr)
library(skimr)
library(kableExtra)
```

# Data

```{r data}
rm(list = ls())
# data <- read_sav("./Data/uncleaned_full_data.sav")

data <- read_sav("../Data/UK23626 Workers sample data with nat rep and graduates weight.sav") %>%
    rename(
    ID = MIProRspId
  ) 

# calculate indicator data
# invert response function for summing
invert_response <- function(x){
  x <- 2 + (-1*x)
}

# Now for just indicators
indicator_data <- data %>% 
  select(ID, starts_with("Q1")) %>%
  reframe(across(starts_with("Q1"), ~invert_response(.x)), .by = ID) %>%
  rowwise() %>%
  mutate(
    sum_true = sum(Q1_1, Q1_2, Q1_3, Q1_4, Q1_5, Q1_6)
  ) %>% 
  select(-starts_with("Q1"))

# Check ID the same
sum(data$ID == indicator_data$ID) == nrow(data)

# Merge into main data set
data <- left_join(data, indicator_data, by = "ID")

data_labelled <- haven::as_factor(data)
```

## Assign people to groups

As of 17 February, the definitions are:

-   **Outsourced**, defined as responding 'I am sure I am outsourced ' or 'I might be outsourced', and responding 'I do work on a long-term basis'.
-   **Likely agency**, defined as those responding 'I am sure I am agency' and 'I do work on a long-term basis', **excluding** those people who are already defined as being outsourced.
-   **High indicators**: defined as responding TRUE to 5 or 6 of the outsourcing indicators, as well as responding 'I do work on a long-term basis', **excluding** those people who are already defined as outsourced or likely agency.

```{r group-assigment}

## These groups are mutually exclusive
data <- data %>%
  mutate(
    # SURE outsourced or MIGHT BE outsourced + LONGTERM
    outsourced = ifelse((Q3v3a == 1 & Q2 == 1) | (Q3v3a == 2 & Q2 == 1), 1, 0),
    # NOT outsourced, SURE agency, and LONG-TERM
    likely_agency = ifelse(outsourced == 0 & Q2 == 1 & (Q3v3b == 1 | Q3v3c == 1 | Q3v3d == 1), 1, 0),
    likely_agency = ifelse(is.na(likely_agency), 0, likely_agency),
    # NOT outsourced, NOT likely agency, 5 or more indicators, & LONGTERM
    high_indicators = ifelse(outsourced == 0 & likely_agency == 0 & (Q2 == 1 & sum_true >= 5), 1, 0)
  )

# count the groupings
lapply(list(data$outsourced,
            data$likely_agency,
            data$high_indicators), sum)

# Flatten these groupings into a single variable

data <- data %>%
  mutate(
    outsourcing_group = factor(case_when(outsourced == 1 ~ 'outsourced',
                                         likely_agency == 1 ~ 'likely_agency',
                                         high_indicators == 1 ~ 'high_indicators',
                                         TRUE ~ 'not_outsourced'), 
                               levels = c("not_outsourced",
                                          "outsourced",
                                          "likely_agency",
                                          "high_indicators")
    )
  )

```

# Demographics

The demographic variables explored here are:

-   Age

-   Sex

-   Ethnicity

-   Region

-   Income

-   Born UK

## Age

```{r demographics, output:true}

age_statistics <- data %>%
  # get values of labels
  mutate_all(haven::as_factor) %>%
  group_by(outsourcing_group) %>%
  mutate(
    # make age numeric, not a labelled factor
    Age = as.numeric(as.character(Age))
  ) %>%
  summarise(
    mean = mean(Age, na.rm = T),
    median = median(Age,na.rm = T),
    min = min(Age,na.rm = T),
    max = max(Age,na.rm = T),
    stdev = sd(Age,na.rm = T)
  )

knitr::kable(age_statistics, 
             digits = 2, 
             col.names = c("Outsourcing group",
                           "Mean",
                           "Median",
                           "Min",
                           "Max",
                           "Standard dev.")) %>%
  kable_styling(full_width = F)
```

## Sex

```{r}

sex_statistics <- data %>%
  # get values of labels
  mutate_all(haven::as_factor) %>%
  group_by(outsourcing_group, Sex) %>%
  summarise(
    Frequency = n()
  ) %>%
  mutate(
    Percentage = 100 * (Frequency / sum(Frequency))
  ) %>%
  rename(
    `Outsourcing group` = outsourcing_group
  )

knitr::kable(sex_statistics, 
             digits = 2) %>%
  kable_styling(full_width = F)
```

## Ethnicity

```{r}
ethnicity_statistics <- data %>%
  # get values of labels
  # mutate_all(haven::as_factor) %>%
  group_by(outsourcing_group, Ethnicity) %>%
  summarise(
    Frequency = n()
  ) %>%
  mutate(
    Percentage = 100 * (Frequency / sum(Frequency))
  ) %>%
  rename(
    `Outsourcing group` = outsourcing_group
  )

ethnicity_statistics %>%
  mutate(
    Ethnicity = haven::as_factor(Ethnicity)
    ) %>%
  knitr::kable(.,digits = 2) %>%
  kable_styling(full_width = F)

ethnicity_statistics %>%
  ggplot(., aes(`Outsourcing group`, Percentage, fill = as.factor(Ethnicity))) +
  geom_col(colour="black") +
  annotate("text",inherit.aes=F, data=ethnicity_statistics, aes(`Outsourcing group`, y = 50, label = sum(ethnicity_statistics$Frequency))) +
  coord_flip()

```

Another way of looking at this is to calculate for each ethnicity, the proportion of workers in each outsourcing group. Doing so yields the plot below

```{r}
ethnicity_statistics <- data %>%
  # get values of labels
  # mutate_all(haven::as_factor) %>%
  group_by(Ethnicity,outsourcing_group) %>%
  summarise(
    Frequency = n()
  ) %>%
  mutate(
    Percentage = 100 * (Frequency / sum(Frequency))
  ) %>%
  rename(
    `Outsourcing group` = outsourcing_group
  )

ethnicity_statistics %>%
  mutate(
    Ethnicity = haven::as_factor(Ethnicity)
    ) %>%
  knitr::kable(.,digits = 2) %>%
  kable_styling(full_width = F)

ethnicity_statistics %>%
  mutate(
    Ethnicity = haven::as_factor(Ethnicity)
    ) %>%
  ggplot(., aes(Ethnicity, Percentage, fill = `Outsourcing group`)) +
  geom_col(colour="black") +
  coord_flip()
```

## Region

```{r}
region_statistics <- data %>%
  # get values of labels
  # mutate_all(haven::as_factor) %>%
  group_by(outsourcing_group, Region) %>%
  summarise(
    Frequency = n()
  ) %>%
  mutate(
    Percentage = 100 * (Frequency / sum(Frequency))
  ) %>%
  rename(
    `Outsourcing group` = outsourcing_group
  )

region_statistics %>%
  mutate(
    Region = haven::as_factor(Region)
    ) %>%
  knitr::kable(.,digits = 2) %>%
  kable_styling(full_width = F)

region_statistics %>%
  mutate(
    Region = haven::as_factor(Region)
    ) %>%
  ggplot(., aes(`Outsourcing group`, Percentage, fill = Region)) +
  geom_col(colour="black") +
  coord_flip()
```

## Income

```{r}
data <- data %>%
  # get values of labels
  # mutate_all(haven::as_factor) %>%
  mutate(
    # make all annual incomes. Note this assumes 52 working weeks!
    income_annual = case_when(INCOME_FREQ == 1 ~ INCOME_OPEN_1,
                              INCOME_FREQ == 2 ~ INCOME_OPEN_1*12,
                              INCOME_FREQ == 3 ~ INCOME_OPEN_1*52,
                              INCOME_FREQ == 4 ~ INCOME_OPEN_1*HOURS*52,
                              TRUE ~ NA)
  ) 

income_statistics <- data %>%
  group_by(outsourcing_group) %>%
  summarise(
    mean = mean(income_annual, na.rm = T),
    median = median(income_annual,na.rm = T),
    min = min(income_annual,na.rm = T),
    max = max(income_annual,na.rm = T),
    stdev = sd(income_annual,na.rm = T)
  )

knitr::kable(income_statistics, 
             digits = 2, 
             col.names = c("Outsourcing group",
                           "Mean",
                           "Median",
                           "Min",
                           "Max",
                           "Standard dev.")) %>%
  kable_styling(full_width = F)

options(scipen = 999)
data %>%
  ggplot(., aes(outsourcing_group, income_annual)) + 
  geom_violin() +
  geom_text(inherit.aes=F, data=income_statistics, aes(outsourcing_group, y = 5e+05), label=paste0("Mean = ", round(income_statistics$mean,0),"\n", "Median = ", income_statistics$median), nudge_x = 0.1, hjust=0) +
  coord_cartesian(xlim=c(1,4.5))
```

Clearly there is some big skew here. The case close to 3million is someone who said they were paid 55000 a week. They are a director of a company, but say they are sure they're outsourced and sure they're an agency worker. This looks to be a poor case and should probably be removed.
